Redshift evolution of the underlying type Ia supernova stretch distribution

### 2021-03-05 Nicolas.tex
Language editor files

\newcommand{\orcid}[1]{
\href{https://orcid.org/#1}{
\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}

@@

\begin{document}
\title{Redshift Evolution of the Underlying Type Ia Supernova Stretch
Distribution}

\titlerunning{Redshift Evolution of the Underlying Type Ia Supernova Stretch
Distribution}
\authorrunning{N.~Nicolas et al.}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties rises, notably that of SN~Ia population evolution. In this
    paper, we study the dependence with redshift of the SN~Ia \texttt{SALT2.4}
    lightcurve stretch, a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to strongly correlate with the
    SN environment, notably with stellar age tracers. We model the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
    and assuming constant underlying stretch distribution for each age
    population consisting of Gaussian mixtures. We test our prediction against
    published samples cut to have marginal magnitude selection effects, so that
    any observed change is indeed astrophysical and not observational in origin.
    In this first study, there are indications that the underlying SN~Ia stretch
    distribution is evolving as a function of redshift, and that the young/old
    drifting model is a better description of the data than any time-constant
    model, including the sample-based asymmetric distributions often used to
    correct Malmquist bias at a significance of over 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode exclusive
    to old environments. The precise impact of the redshift evolution of the SN
    Ia population intrinsic properties on cosmology remains to be studied. Yet,
    the astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections and thereby distances derived from SN affected by
    observational selection effects. We highlight that such a bias will increase
with surveys covering increasingly larger redshift ranges, which is particularly
important for LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant ($H_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ derived when
the SNe~Ia are anchored to Cepheids \citep[the SH0ES
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seem to favor an intermediate value of
$H_0$ \citep{freedman2019, freedman2020}, time delay measurements from strong
lensing seem to also favor high $H_0$ values \citep{wong2019}.

The $H_0$ tension has received a lot of attention, as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this $H_0$
tension when accounting for all other probes \citep{knox2019}. Alternatively,
systematic effects affecting one or several of the aforementioned analyses
might explain at least some of this tension. \cite{rigault2015}
suggested that SNe Ia from the Cepheid calibrator sample differ by construction
from those in the Hubble flow sample, as the former strongly favors
young stellar populations, while the latter does not. This selection
effect would impact the derivation of $H_0$ if SNe~Ia from young and older
environments differ in average standardized magnitudes.

The relation between SNe~Ia and their host galaxy properties has been studied
extensively. The first key finding was that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
host galaxies being brighter on average \cite[e.g.][]{kelly2010, sullivan2010,
childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
currently used in cosmological analyses \citep[e.g.][]{betoule2014,
scolnic2018a}, including for deriving $H_0$ \citep{riess2016, riess2019}. Yet,
the underlying connection between the SNe and their host galaxies remains
unclear, particularly when using global properties such as the host stellar
mass. This in turn raises the question of the accuracy of such corrections since
host galaxy global properties evolve with redshift. More recently, studies have
used the local SN host galaxy environment to probe more direct connections
between the SNe and their host galaxy environments \citep{rigault2013}, showing
that local age tracers such as the Local specific Star Formation Rate (LsSFR) or
the local color are more strongly correlated with the standardized SN magnitude
\citep{roman2018, kim18, rigault2020}. Further reinforcing this connection is
the identification of SN~Ia spectral features that are correlated with LsSFR
\citep{nordin2018}. These results suggest age as the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology, since the environmental correction to apply to SN standardization
could strongly vary with redshift \citep{rigault2013, childress2014,
scolnic2018a}. Yet, the importance of local SN environmental studies remains
highly debated \cite[e.g.][]{jones2015, jones2019} and especially the impact of
such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called ``prompt and delayed'' or ``A+B''
model). \cite{rigault2020} used the LsSFR to classify which are the younger
(those with a high LsSFR) and which are the older (those with low LsSFR).
Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
hamuy2000}. This correlation that has been extensively confirmed since
\citep[e.g.][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
Following the ``A+B'' model and the connection between SN stretch and host
galaxy properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this question
using the most recent SNe~Ia datasets.

In this paper, we take a step aside from the cosmological analyses to probe the
validity of our modeling of the SN population, which we claim to be constituted
of two age-populations \citep{rigault2013, rigault2015, rigault2020}: one old
and one younger, the former having on average lower lightcurve stretches and
being brighter after standardization. We use the correlation between the SN age,
as probed by the LsSFR, and the SN stretch to model the expected evolution of
the underlying SN stretch distribution as a function of redshift. This modeling
relies on three assumptions: (1) there are two distinct populations of SNe~Ia;
(2) the relative fraction of each of these populations as a function of redshift
follows the model presented in \cite{rigault2020} and (3) the underlying
distribution of stretch for each age sample is constant. This paper aims at
testing this specific model with datasets from the literature.

We present in Section~\ref{sec:sample} the sample we are using for this
analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
importance of obtaining a ``complete'' sample, i.e.\ representative of the true
underlying SNe Ia distribution, and how we build one from the Pantheon sample.
We then present in Section~\ref{sec:modeling} our modeling of the distribution
of stretch and our results are presented in Section~\ref{sec:results}. In this
section, we test whether the SN stretch distribution evolves as a function of
redshift and if the aforementioned age model is in good agreement with this
evolution. We briefly discuss these results in the context of SN cosmology in
Section~\ref{sec:discussion} and we conclude in Section~\ref{sec:ccl}.

\section{Complete Sample Construction}\label{sec:sample}

The ideal SN~Ia sample for studying this question would be a very deep,
large-area, volume-limited sample. This would capture the true underlying
stretch distribution function, and we would then study how it evolves with
redshift. No such sample exists, so first we must construct subsamples that are
as near to volume-limited as possible from existing high-redshift SN~Ia samples.

@@

We base our analysis on the most recent comprehensive SNe~Ia compilation, the
Pantheon catalog from \cite{scolnic2018a}. A naive approach to test the SN
stretch redshift drift would be to simply compare the observed SN stretch
distributions in a few bins of redshift. In practice, however, differential
observational selection effects will affect the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SNe~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if magnitude-related observational selection effects are
not accounted for, one might confuse true population drift with survey
properties, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
SN types and host galaxy redshifts, the observation selection
effects of magnitude-limited surveys should be negligible below a given redshift
at which even the faintest normal SNe~Ia can be observed.
Aiding in the construction of nearly volume-limited subsamples is the
    fact that the SN~Ia population trails off towards fainter SNe~Ia. A
    complication is that complete spectroscopic follow-up has not always been
    the norm, as discussed below. In contrast, targeted surveys have highly
    complex observational selection functions and so are
    discarded from our analysis. High-redshift SN cosmology samples,
    such as Pantheon, are predominately from magnitude-limited surveys from
which volume-limited SN~Ia subsamples can be constructed.

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007, betoule2014}; see also
\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
x_1 + \beta c$ along the aforementioned ellipse given the standardization
coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the
faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. Since one
ought to detect this object typically 5 days before and a week after peak to
build a suitable lightcurve, the effective limiting standardized absolute
magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
limit $m_{\lim}$ of a magnitude limited survey, one can derive the maximum
redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
relation between apparent magnitude, redshift and absolute magnitude
$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
        the Pantheon catalog. The individual SNe are shown as blue dots. The
        ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. The gray diagonal lines
        represent the $(x_1, c)$ evolution for $m = m_{\lim}$, for $z$ between
        $0.50$ and $z=1.70$ using SNLS's $m_{\lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

We will thus consider a set of cuts that will define a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. However, as this solution might be an overly
    simplified way to create a complete sample, e.g., because it ignores
    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
    also consider another set of cuts to define a so-called ``conservative''
    sample. The latter is smaller and therefore less statistically constraining,
    but also even less prone to observational selection effects. If the redshift
    drift is still significant in the conservative sample, it would be even more
    meaningful in a carefully-tailored selection-free sample. These samples are
    adequate for the goal of this study, which is to develop a first
    implementation of a model for drift in SN Ia properties. If fruitful, the
    sample selection can later be refined, if needed, with a more detailed model
of the observational selection, e.g., using the SNANA package \citep{SNANA}.

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, which has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative samples, respectively, for SNLS.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative samples, respectively, for PS1.

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys had to contend with limited spectroscopic resources. As
discussed in \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$.
Based on this, we will use $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative samples, respectively, for SDSS.

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits are roughly located slightly
    before the peak of these histograms. In Section \ref{ssec:verify} we
validate that these redshift limits are effective for constructing nearly
volume-limited subsamples from samples that were initially more closely
magnitude-limited in their search or spectroscopic follow-up.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
        SDSS, PS1 and SNLS dataset respectively (data from Pantheon,
        \citealt{scolnic2018a}). The colored parts represent the distribution of
        SNe~Ia kept in our analysis for they are supposedly free from
    observational selection bias (see Section~\ref{sec:sample}). The
darker (resp.\ lighter) color responds to the conservative (resp.\ fiducial)
selection cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2 ; see \citealt{aldering2020}). For this dataset, spectroscopic
screening was done for candidates with $r \lesssim 19.5$; redshift cuts were
then applied when selecting which SN Ia to follow, resulting in a redshift range
of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} for
extracting local host properties. These 114 SNfactory SNe~Ia are thus in the
volume-limited part of the survey (Aldering et al., in prep.), and are therefore
assumed to be a random sampling of the underlying SN population. The SNfactory
sample is particularly useful for studying SN property drift, as it enables us
to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
sample from Pantheon \citep{strolger04}. These high-redshift SNe are of great
interest as they provide the greatest leverage for testing evolution. While at
these redshifts the supernovae typing is challenging, the target classification
was robust enough to include them within the cosmological analysis
\citep{scolnic2018a} and we do not impose further cuts.
Section~\ref{sec:results} highlights that, while compatible with it, our results
are not dependent on the inclusion of this dataset.

>>>>

\newcommand{\LEt}[1]{\textcolor{red}{\textbf{ #1}}}
\newcommand{\orcid}[1]{
\href{https://orcid.org/#1}{
\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}

@@

\begin{document}

\title{Redshift evolution of the underlying type Ia supernova stretch
distribution}

\titlerunning{Redshift evolution of the underlying type Ia supernova stretch
distribution}
\authorrunning{N.~Nicolas et al.}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties arises, notably that of the evolution of SN~Ia populations. We study the dependence on redshift of the SN~Ia \texttt{SALT2.4}
    light-curve stretch, which is a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to be strongly correlated with the
    SN environment, notably with stellar age tracers. We modeled the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
    and assuming a constant underlying stretch distribution for each age
    population consisting of Gaussian mixtures. We tested our prediction against
    published samples that were cut to have marginal magnitude selection effects, so that
    any observed change is indeed astrophysical and not observational in origin.
    In this first study, there are indications that the underlying SN~Ia stretch
    distribution evolves as a function of redshift, and that the young/old\LEt{the slash commonly denotes "ratio", as in "S/N" for "signal-to-noise ratio". The way you use it here is ambiguous: do you mean "and" or "or", or maybe a range?\ Please rephrase for clarity}
    drifting model is a better description of the data than any time-constant
    model, including the sample-based asymmetric distributions that are often used to
    correct Malmquist bias at a significance higher than 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode that is exclusive
    to old environments. The precise effect of the redshift evolution of the intrinsic properties of a SN
    Ia population on cosmology remains to be studied. The astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections, however, and thereby the distances that are derived based on SN that are affected by
    observational selection effects. We highlight that this bias will increase
with surveys covering increasingly larger redshift ranges, which is particularly
important for the \LEt{please spell this out here and in the main text, you use the abbreviation only one time each}LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery that the Universe expands
\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today for
understanding the properties of dark energy (DE) because it\LEt{grammatically, "it" here refers to "DE". Is this correct, or should this be "they" because you still refer to the SNe Ia, or do you mean "the probe"? If you mean the probe, please rephrase to "because this probe" for clarity} is the only tool that can
precisely map the recent expansion rate ($z<0.5$) when DE is driving it
\citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the key to directly measuring the
Hubble constant ($H_0$) if their absolute magnitude can be calibrated
\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ tha tis derived when
the SNe~Ia are anchored to Cepheids \citep[the\LEt{please provide the spelled-out versions of all instruments and surveys at first occurrence. If this interrupts the flow of the sentence too much, it may be done as a footnote} SH0ES
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seems to favor an intermediate value of
$H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
lensing also appear to favor high $H_0$ values \citep{wong2019}.

The $H_0$ conundrum has received much attention because it could be a sign of new
fundamental physics. No simple solution is so far able to accommodate this $H_0$
conundrum when all other probes are accounted for, however \citep{knox2019}. Alternatively,
systematic effects affecting one or several of the aforementioned analyses
might explain at least some of this discrepancy. \cite{rigault2015}
suggested that SNe Ia from the Cepheid calibrator sample differ by construction
from those in the Hubble flow sample, as the former strongly favors
young stellar populations while the latter does not. This selection
effect would affect the derivation of $H_0$ if SNe~Ia from young and older
environments differed in average standardized magnitudes.

The relation between SNe~Ia and their host galaxy properties has been studied
extensively. The first key finding was that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
host galaxies being brighter on average \cite[e.g.,][]{kelly2010, sullivan2010,
childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
currently used in cosmological analyses \citep[e.g.,][]{betoule2014,
scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The underlying connection between the SNe and their host galaxies remains
unclear, however, particularly when global properties such as the host stellar
mass are used. This in turn raises the question of the accuracy of such corrections because
the global properties of the host galaxies evolve with redshift. More recently, studies have
used the local SN host galaxy environment to probe more direct connections
between the SNe and their host galaxy environments \citep{rigault2013}, showing
that local age tracers such as the local specific star formation rate (LsSFR) or
the local color are more strongly correlated with the standardized SN magnitude
\citep{roman2018, kim18, rigault2020}. The identification of SN~Ia spectral features that are correlated with LsSFR
\citep{nordin2018} further support this connection. These results suggest age as the driving parameter
underlying the mass step. If this is true, itwould have a significant effect for
cosmology because the environmental correction that would need to be applied to SN standardization
could strongly vary with redshift \citep{rigault2013, childress2014,
scolnic2018a}. The importance of local SN environmental studies remains
highly debated, however \cite[e.g.,][]{jones2015, jones2019}, and especially the effect of
such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.\LEt{to properly include these references into the main text, please separate them by commas and set off the last reference additionally with "and". This is a LaTeX command error that I cannot fix for you with the program I work with (citet and citep)}
\cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, a young population that follows the host star formation activity, and an old population that
followed the host stellar mass (the so-called prompt-and-delayed or A+B
model). \cite{rigault2020} used the LsSFR to determine the younger
(those with a high LsSFR) and the older (those with a low LsSFR) populations.
Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
hamuy2000}. This correlation has been extensively confirmed since then
\citep[e.g.,][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
Following the A+B model and the connection between SN stretch and host
galaxy properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this question
using the most recent SNe~Ia datasets.

We here step aside from the cosmological analyses to probe the
validity of our modeling of the SN population, which we claim is constituted
of two age populations \citep{rigault2013, rigault2015, rigault2020}: an older
and a younger population, the former having lower light-curve stretches on average and
being brighter after standardization. We use the correlation between the SN age
as probed by the LsSFR and the SN stretch to model the expected evolution of
the underlying SN stretch distribution as a function of redshift. This modeling
relies on three assumptions: (1) There are two distinct populations of SNe~Ia,
(2) the relative fraction of each of these populations as a function of redshift
follows the model presented in \cite{rigault2020}, and (3) the underlying
distribution of stretch for each age sample is constant. This paper tests this specific model with datasets from the literature.

We present the sample we used for this analysis in Section~\ref{sec:sample}. It was derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
importance of obtaining a ``complete'' sample, that is, a sample that is\ representative of the true
underlying SNe Ia distribution, and how we built one from the Pantheon sample.
We then present our modeling of the distribution
of stretch in Section~\ref{sec:modeling}  and our results in Section~\ref{sec:results}. In this
section, we test whether the SN stretch distribution evolves as a function of
redshift and determine whether the aforementioned age model agrees well with this
evolution. We briefly discuss these results in the context of SN cosmology in
Section~\ref{sec:discussion}, and we conclude in Section~\ref{sec:ccl}.

\section{Complete sample construction}\label{sec:sample}

The ideal SN~Ia sample for studying this question would be a very deep,
large-area, volume-limited sample. This would capture the true underlying
stretch distribution function, and we would then study how it evolves with
redshift. No such sample exists, therefore we must first construct subsamples from existing high-redshift SN~Ia samples that are
as near to volume-limited as possible.

@@

We based our analysis on the most recent comprehensive SNe~Ia compilation, the
Pantheon catalog from \cite{scolnic2018a}. A naive approach to testing the SN
stretch redshift drift would be to simply compare the observed SN stretch
distributions in a few redshift bins. In practice, however, differential
observational selection effects will affect the observed SN
stretch distributions. Because the observed SN~Ia magnitude correlates
with the light-curve stretch (and color), the first SNe~Ia that a
magnitude-limited survey will miss are those at the lowest stretch (which are the reddest).
Consequently, if magnitude-related observational selection effects are
not accounted for, true population drift might be confused with survey
properties, and vice versa.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring
SN types and host galaxy redshifts, the observation selection
effects of magnitude-limited surveys should be negligible below a given redshift
at which even the faintest normal SNe~Ia can be observed.
Aiding in the construction of nearly volume-limited subsamples is the
    fact that the SN~Ia population trails off toward fainter SNe~Ia. A
    complication is that complete spectroscopic follow-up has not always been
    the norm, as discussed below. In contrast, targeted surveys have highly
    complex observational selection functions and so are
    discarded from our analysis. High-redshift SN cosmology samples,
    such as the Pantheon sample, are predominately assembled from magnitude-limited surveys from
which volume-limited SN~Ia subsamples can be constructed.

We present in Fig.~\ref{fig:maglim} the light-curve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008}, and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007, betoule2014}; see also
\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
x_1 + \beta c$ within the aforementioned ellipse given the standardization
coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The
faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought to be detected 5 days before and a week after peak to
build a suitable light curve, the effective limiting standardized absolute
magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
limit $m_{\lim}$ of a magnitude limited survey, we can derive the maximum
redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
relation between apparent magnitude, redshift, and absolute magnitude
$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        light-curve parameters of SNe~Ia from the SDSS, PS1, and SNLS samples of
        the Pantheon catalog. The individual SNe are shown as blue dots. The
        ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. The diagonal gray lines
        represent the $(x_1, c)$ evolution for $m = m_{\lim}$ for $z$ between
        $0.50$ and $z=1.70$ using the SNLS $m_{\lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

We therefore considered a set of cuts that defines a first fiducial
    sample, taking the limits as initially suggested by the previous
    completeness analysis. However, as this solution might be an overly
    simplified way to create a complete sample, for example, because it ignores
    spectroscopic follow-up in efficiency for redshifts below $z_{lim}$, we
    also considered another set of cuts to define a so-called conservative
    sample. This is smaller and therefore less statistically constraining,
    but also even less prone to observational selection effects. If the redshift
    drift is still significant in the conservative sample, it would be even more
    meaningful in a carefully tailored selection-free sample. These samples are
    adequate for the goal of this study, which is to develop a first
    implementation of a model for drift in SN Ia properties. If fruitful, the
    sample selection can later be refined as required with a more detailed model
of the observational selection, for instance, using the SNANA package \citep{SNANA}.

The SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, which has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts into a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010}, and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010}, however, suggests a lower limit of
$z_{\lim}=0.55$. We therefore used $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative samples, respectively, for the SNLS.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields $z_{\lim}=0.31$, in
agreement with Fig.~6 of \cite{scolnic2018a}, for example. If we were to be
    conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; we therefore used $0.31$ and $0.27$ for our fiducial and
conservative samples, respectively, for PS1.

In a similar redshift range, the SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to $z_{\lim}=0.24$. However, the
SDSS surveys had to contend with limited spectroscopic resources. As
discussed in \citet[][Section~2]{kessler2009}, during the first year of SDSS,
SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up, corresponding
to a redshift cut at $0.15$. For the remaining SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} showed a reasonable completeness up to $z_{\lim}=0.2$.
Based on this, we used $z_{\lim}=0.20$ and $z_{\lim}=0.15$ for our
fiducial and conservative samples, respectively, for the SDSS.

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits are roughly located slightly
    before the peak of these histograms. In Section \ref{ssec:verify} we
confirm that these redshift limits are effective for constructing nearly
volume-limited subsamples from samples that were initially more closely
magnitude limited in their search or spectroscopic follow-up.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia from the
        SDSS, PS1, and SNLS dataset (data from Pantheon,
        \citealt{scolnic2018a}). The colored parts represent the distribution of
        SNe~Ia that we retained in our analysis because they are assumed to be free from
    observational selection bias (see Section~\ref{sec:sample}). The
darker (lighter) color responds to the conservative (fiducial)
selection cut.}
    \label{fig:cuts}
\end{figure}

In addition, we used the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2020} and that have
been discovered from nontargeted searches (114 SNe~Ia, see their sections~3
and~4.2.2 ; see \citealt{aldering2020}). For this dataset, spectroscopic
screening was performed for candidates with $r \lesssim 19.5$; redshift cuts were
then applied when selecting which SN Ia to follow, resulting in a redshift range
of $0.02 < z < 0.09$, further reduced to $<0.08$ in \cite{rigault2020} to extract local host properties. These 114 SNfactory SNe~Ia are thus in the
volume-limited part of the survey (Aldering et al., in prep.), and are therefore
assumed to be a random sampling of the underlying SN population. The SNfactory
sample is particularly useful for studying SN property drift because it enables us
to have a large complete SN~Ia sample at $z<0.1$. Finally, we include the HST
sample from Pantheon \citep{strolger04}. These high-redshift SNe are of great
interest as they provide the greatest leverage for testing evolution. While at
these redshifts the supernovae typing is challenging, the target classification
was robust enough to include them in the cosmological analysis
\citep{scolnic2018a}, and we did not impose further cuts.
Section~\ref{sec:results} highlights that while compatible with it, our results
are not dependent on the inclusion of this dataset.

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. We observe here that
the fraction of low-stretch SNe (typically $x_1 < -1$) appears to decrease as a
function of redshift; this is confirmed in Fig.~\ref{fig:modelall}, in which the
evolution of the mean stretch is shown, with the data split in redshift bins of
regular sample size. SNe~Ia at higher redshift have a
larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) on average than those at lower redshift
($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
distribution evolves with redshift.
